A weakly structured stem for human origins in Africa

### 2022-07-22 paper.tex
Revised "ecological riddle" paragraph

Evidence for archaic hominin admixture in Eurasia has bolstered the plausibility of
archaic hominin admixture having also occurred in Africa. For this reason, previous
work has focused on archaic hominin admixture to explain patterns of polymorphism
inconsistent with a single-origin model. Here, we have shown that
weakly-structured-stem models better capture these patterns.
They also help explain an
ecological riddle posed by the archaic hominin admixture model. Neanderthal populations
were separated from early \emph{Homo sapiens} by thousands of kilometers and
continental geographic barriers. By contrast, an archaic hominin population in
Africa would need to have stayed in relative reproductive isolation from the
ancestral human lineage over hundreds of thousands of years despite closer geographic
proximity and reproductive compatibility. The weakly-structured-stem model
resolves this ecological riddle by accommodating continuous or recurrent
contact between two or more groups present in Africa.

>>>>

Evidence for archaic hominin admixture in Eurasia has bolstered the plausibility of 
archaic hominin admixture having also occurred in Africa. For this reason, previous 
work has focused on archaic hominin admixture to explain patterns of polymorphism 
inconsistent with a single-origin model. Here, we have shown that weakly-structured-stem 
models better capture these patterns. Preferring weakly-structured-stem models over 
archaic-admixture models potentially has ecological implications. With a weakly 
structured stem, there is no need to posit that an archaic hominin population stayed 
reproductively isolated from the ancestral human lineage for thousands of years before 
the initiation of gene flow. Instead, there would have been continuous or recurrent 
contact between two or more groups present in Africa. It would be fruitful to 
investigate the ecological reasons for continued contact between groups within Africa 
but long periods of isolation between African and Eurasian groups.
